---
title: "Data wrangling for Class Project"
author: "Maggie Douglas"
date: "`r format(Sys.time(), '%m/%d/%y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Purpose

This code is meant to match PPL data to building information and reorganize it to generate electricity data by building (or combinations of buildings for those metered together). 

The main steps involved are:

+ Read in the PPL electricity data, building information data, and a key to connect them by name
    + Create and store an annual summary of the electricity data by meter
+ Restructure the building data to generate a single row for the Main Meter and Weis Meter buildings
+ Join PPL electricity data (daily and annual) to building information using a key that connects the two by building name
+ Aggregate electricity and square footage data by building 
+ Inspect the data using a table
+ Generate a few summary graphs 
+ Export the cleaned and reorganized data for downstream analyses

## Load libraries + data

```{r load, message = F}
library(tidyverse) # load tidyverse
```

```{r load_data, message = F, results = 'hide'}
# load electricity data + reformat date
kwh <- read.csv("../data/FY25 PPL Electricity Data.csv", strip.white = T) %>%
  mutate(date = mdy(date)) %>%
  complete(nesting(account_number, meter_origin, meter_number), date) %>%
  mutate(period = ifelse((date > ymd("2024-08-28") & date < ymd("2024-12-21")), "Fall Semester", 
                         ifelse((date > ymd("2025-01-07") & date < ymd("2025-05-18")), "Spring Semester",
                                ifelse(date > ymd("2024-12-21") & date < ymd("2025-01-07"), "Winter Break", "Summer Break")))) 
  
kwh_sub <- read.csv("../output/kwh_main_daily.csv", strip.white = T) %>%
  mutate(date = ymd(date)) %>%
  mutate(period = ifelse((date > ymd("2024-08-28") & date < ymd("2024-12-21")), "Fall Semester", 
                         ifelse((date > ymd("2025-01-07") & date < ymd("2025-05-18")), "Spring Semester",
                                ifelse(date > ymd("2024-12-21") & date < ymd("2025-01-07"), "Winter Break", "Summer Break")))) 

# store temp data for later
temp <- dplyr::select(kwh, date, ave_temp) %>%
  unique()
  
# load building data + clean up building names
buildings <- read.csv("../keys/fy25_building_list_updated.csv", 
                      strip.white = T) %>%
  mutate(NAME = str_remove_all(NAME, "/"),
         NAME = str_replace_all(NAME, "  "," "),
         NAME = str_replace_all(NAME, "   "," "))

# load keys to link datasets
key <- read.csv("../keys/meter_building_key.csv", strip.white = T) %>%
  mutate(NAME = str_remove_all(NAME, "/"),
         NAME = str_replace_all(NAME, "  "," "),
         NAME = str_replace_all(NAME, "   "," "))

key_sub <- read.csv("../keys/submeter_building_key.csv", strip.white = T) %>%
  mutate(NAME = str_remove_all(NAME, "/"),
         NAME = str_replace_all(NAME, "  "," "),
         NAME = str_replace_all(NAME, "   "," ")) 

key_occ <- read.csv("../keys/occupancy_key.csv", 
                      strip.white = T) %>%
  mutate(NAME = str_remove_all(NAME, "/"),
         NAME = str_replace_all(NAME, "  "," "),
         NAME = str_replace_all(NAME, "   "," "))

# load occupancy data + generate mean for AY
occupants <- read.csv("../data/housing_counts.csv",
                      strip.white = T) %>%
  mutate(sem_yr = paste(semester, year)) %>% # create semester column
  filter(sem_yr %in% c("Fall 2024", "Spring 2025")) %>% # filter to FY25
  group_by(building_name) %>% 
  summarize(occupants = mean(occupants, na.rm = T)) %>% # generate mean occupants across semester
  left_join(key_occ, by = "building_name") %>% # join to building names
  group_by(NAME) %>%
  summarize(occupants = sum(occupants, na.rm = T)) # sum across buildings with multiple rows

# PPL data
## store annual totals for kWh
## checked - buildings with missing days do not have weirdly high values when they come back online
kwh_annual <- kwh %>%
  filter(!is.na(total_kwh)) %>%
  group_by(meter_origin) %>%
  summarize(kwh = sum(total_kwh, na.rm = T),
            days_perc = (n()/365)*100)

kwh_academic <- kwh %>%
  filter(period %in% c("Spring Semester","Fall Semester")) %>%
  filter(!is.na(total_kwh)) %>%
  group_by(meter_origin) %>%
  summarize(kwh = sum(total_kwh, na.rm = T),
            days_perc = (n()/244)*100) # days in academic year only

## Submeter data
# buildings with missing days are variable in how they come back online
# Library has a strange high value on 2/17
# Kline and CHW don't have obvious jump after missing data
# all other buildings (93% of days) have strange values 9/02 & 9/06
kwh_sub_clean <- kwh_sub %>%
  mutate(kwh = ifelse(kwh == 0, NA, kwh)) %>% # fill in NA for zero values
  mutate(kwh = ifelse(building == "Library" & date == ymd("2025-02-17"), NA, kwh)) %>% # replace outlier w/ NA
  mutate(kwh = ifelse(date %in% c(ymd("2024-09-02"),ymd("2024-09-06")), NA, kwh)) %>%
  filter(! building %in% c("Rector","HUB")) # these two buildings have < 10% of days present

kwh_sub_annual <- kwh_sub_clean %>%
  filter(!is.na(kwh)) %>%  # filter out NAs
  group_by(building) %>%
  summarize(kwh = sum(kwh, na.rm = T),
            days_perc = (n()/365)*100) %>% # counts percent of days w/ non-zero data
  filter(! building %in% c("Rector","HUB")) # these two buildings have < 10% of days present

kwh_sub_academic <- kwh_sub_clean %>%
  filter(period %in% c("Spring Semester","Fall Semester")) %>%
  filter(!is.na(kwh)) %>%  # filter out NAs
  group_by(building) %>%
  summarize(kwh = sum(kwh, na.rm = T),
            days_perc = (n()/244)*100) %>% # counts percent of days w/ non-zero data
  filter(! building %in% c("Rector","HUB")) # these two buildings have < 10% of days present
```

## Check key data

```{r check, results = 'hide'}
str(kwh)
str(buildings)
str(occupants)
```

## Restructure building data 

+ Generate total square footage for buildings on the Main Meter and fill in values for other variables
+ Similar operation for buildings on the Weis Meter
+ Filter building info to those with individual meters and then add back in the Main Meter and Weis aggregated information

```{r restructure}
# Generate lookup table for each meter status

buildings_main <- buildings %>%
  filter(main_meter == 1) %>% # filter to main meter buildings
  summarize(sqft = sum(sqft, na.rm = T)) %>% # sum sqft for these buildings
  mutate(meter = "Main Meter - Total",
         NAME = "Main Meter",
         type_new = "Main Meter") %>%
  select(type_new, NAME, sqft, meter)

buildings_weis <- buildings %>%
  filter(weis_meter == 1) %>% # filter to weis buildings
  summarize(sqft = sum(sqft, na.rm = T)) %>%
    mutate(meter = "Weis Meter - Total",
           NAME = "Weis Meter",
         type_new = "Weis Meter") %>%
    select(type_new, NAME, sqft, meter)

buildings_agg <- rbind(buildings_main, buildings_weis)

buildings_individual <- buildings %>%
  filter(main_meter == 0 & weis_meter == 0) %>% # keep only buildings on individual meters
  mutate(meter = "Individual") %>%
  select(meter, NAME, type_new, occupant, address, date_constr, date_acqd, date_reno, sqft, rental) %>% # select relevant columns
 # rbind(buildings_main, buildings_weis) %>%
  select(type_new, NAME, sqft, meter) 

buildings_submeter <- buildings %>%
  filter(main_disagg ==1) %>% # keep only buildings on individual meters
  mutate(meter = "Submeter") %>%
  select(meter, NAME, type_new, occupant, address, date_constr, date_acqd, date_reno, sqft, rental) %>% # select relevant columns
 # rbind(buildings_main, buildings_weis) %>%
  select(type_new, NAME, sqft, meter)

# Store summary of building meter status
buildings_sum <- buildings %>%
    mutate(meter = ifelse(weis_meter == 1, "Weis Meter", 
                        ifelse(main_meter == 1 & main_disagg == 0, "Main Meter", 
                               ifelse(main_disagg == 1, "Submeter", "Individual"))))
```

## Wrangle datasets

### Annual totals

```{r annual, warning = F, message = F}
# generate annual summary for individually metered buildings
joined_individual <- kwh_annual %>%
  left_join(key, by = "meter_origin") %>% # join to key to match meters to building names
  right_join(buildings_individual, by = "NAME", relationship = "many-to-one") %>% # join to building info by name
  group_by(type_new, NAME, meter) %>% # group by building
  summarize(days_perc = min(days_perc, na.rm = T), # take minimum days_perc to reflect coverage of whole data
            kwh = sum(kwh, na.rm = T), # sum kwh by building
            sqft = mean(sqft, na.rm = T)) %>% # take mean to preserve sqft data
  mutate(type = type_new) %>%
  left_join(occupants, by = "NAME") %>%
  ungroup() %>%
  mutate_all(~ifelse(is.nan(.), NA, .)) %>%
  select(type, meter, NAME, days_perc, kwh, sqft, occupants)

# generate annual summary for submetered buildings
joined_sub <- kwh_sub_annual %>%
  left_join(key_sub, by = "building") %>%
  left_join(buildings_submeter, by = "NAME", relationship = "many-to-one") %>% # join to building info by name
  filter(building != "CHW_Base") %>% # remove duplicate CHW value - not sure what this is?
  group_by(type_new, NAME, days_perc, meter) %>% # group by building
  summarize(kwh = sum(kwh, na.rm = T), # sum kwh by building
            sqft = mean(sqft, na.rm = T)) %>% # preserve sqft as is
  mutate(type = type_new) %>%
 # filter(!is.na(type_new)) %>% # filter out those with NA for type
  left_join(occupants, by = "NAME") %>%
  ungroup() %>%
  mutate_all(~ifelse(is.nan(.), NA, .)) %>%
  select(type, meter, NAME, days_perc, kwh, sqft, occupants)

# generate annual summary for main and weis meters
joined_agg <- kwh_annual %>%
  left_join(key, by = "meter_origin") %>%
  right_join(buildings_agg, by = "NAME") %>%
  group_by(type_new, NAME, days_perc, meter) %>% # group by building
  summarize(kwh = sum(kwh, na.rm = T), # sum kwh by building
            sqft = mean(sqft, na.rm = T)) %>% # preserve sqft as is
  mutate(type = type_new) %>%
 # filter(!is.na(type_new)) %>% # filter out those with NA for type
  left_join(occupants, by = "NAME") %>%
  ungroup() %>%
  mutate_all(~ifelse(is.nan(.), NA, .)) %>%
  select(type, meter, NAME, days_perc, kwh, sqft, occupants)

# combine into one data frame
joined_full <- rbind(joined_individual, joined_sub, joined_agg) %>%
  filter(kwh != 0) %>%
  mutate(kwh_corr = kwh/(days_perc/100)) %>% # estimate kwh for all days in the year
  arrange(type, meter) %>%
  select(type, meter, NAME, days_perc, kwh, kwh_corr, sqft, occupants) 

# store lookup for data coverage
data_cov <- joined_full %>%
  select(NAME, days_perc)
```

### Academic totals

```{r academic, warning = F, message = F}
# generate annual summary for individually metered buildings
joined_ind_aca <- kwh_academic %>%
  left_join(key, by = "meter_origin") %>% # join to key to match meters to building names
  right_join(buildings_individual, by = "NAME", relationship = "many-to-one") %>% # join to building info by name
  group_by(type_new, NAME, meter) %>% # group by building
  summarize(days_perc = min(days_perc, na.rm = T), # take minimum days_perc to reflect coverage of whole data
            kwh = sum(kwh, na.rm = T), # sum kwh by building
            sqft = mean(sqft, na.rm = T)) %>% # take mean to preserve sqft data
  mutate(type = type_new) %>%
  left_join(occupants, by = "NAME") %>%
  ungroup() %>%
  mutate_all(~ifelse(is.nan(.), NA, .)) %>%
  select(type, meter, NAME, days_perc, kwh, sqft, occupants)

# generate annual summary for submetered buildings
joined_sub_aca <- kwh_sub_academic %>%
  left_join(key_sub, by = "building") %>%
  left_join(buildings_submeter, by = "NAME", relationship = "many-to-one") %>% # join to building info by name
  filter(building != "CHW_Base") %>% # remove duplicate CHW value - not sure what this is?
  group_by(type_new, NAME, days_perc, meter) %>% # group by building
  summarize(kwh = sum(kwh, na.rm = T), # sum kwh by building
            sqft = mean(sqft, na.rm = T)) %>% # preserve sqft as is
  mutate(type = type_new) %>%
 # filter(!is.na(type_new)) %>% # filter out those with NA for type
  left_join(occupants, by = "NAME") %>%
  ungroup() %>%
  mutate_all(~ifelse(is.nan(.), NA, .)) %>%
  select(type, meter, NAME, days_perc, kwh, sqft, occupants)

# generate annual summary for main and weis meters
joined_agg_aca <- kwh_academic %>%
  left_join(key, by = "meter_origin") %>%
  right_join(buildings_agg, by = "NAME") %>%
  group_by(type_new, NAME, days_perc, meter) %>% # group by building
  summarize(kwh = sum(kwh, na.rm = T), # sum kwh by building
            sqft = mean(sqft, na.rm = T)) %>% # preserve sqft as is
  mutate(type = type_new) %>%
 # filter(!is.na(type_new)) %>% # filter out those with NA for type
  left_join(occupants, by = "NAME") %>%
  ungroup() %>%
  mutate_all(~ifelse(is.nan(.), NA, .)) %>%
  select(type, meter, NAME, days_perc, kwh, sqft, occupants)

# combine into one data frame
joined_full_aca <- rbind(joined_ind_aca, joined_sub_aca, joined_agg_aca) %>%
  filter(kwh != 0) %>%
  mutate(kwh_corr = kwh/(days_perc/100)) %>% # estimate kwh for all days in the year
  arrange(type, meter) %>%
  select(type, meter, NAME, days_perc, kwh, kwh_corr, sqft, occupants) 

# store lookup for data coverage
data_cov_aca <- joined_full_aca %>%
  select(NAME, days_perc)
```

### Daily data

```{r split, warning = F, message = F}
# generate daily summary for individually metered buildings
daily_individual <- kwh %>%
  left_join(key, by = "meter_origin") %>% # join to key to match meters to building names
  right_join(buildings_individual, by = "NAME", relationship = "many-to-many") %>% # join to building info by name
  group_by(type_new, NAME, date, meter) %>% # group by building
  summarize(kwh = sum(total_kwh, na.rm = T), # sum kwh by building
            sqft = mean(sqft, na.rm = T)) %>%
  mutate(type = type_new) %>%
  ungroup() %>%
  mutate_all(~ifelse(is.nan(.), NA, .)) %>% # convert NaN to NA
  select(type, meter, date, NAME, kwh, sqft) %>%
  complete(nesting(type, meter, NAME), date) %>% # restore NA values for missing date-building combos
  filter(!is.na(date)) # remove rows without value for date

# generate daily summary for submetered buildings
daily_sub <- kwh_sub_clean %>%
  left_join(key_sub, by = "building") %>%
  left_join(buildings_submeter, by = "NAME", relationship = "many-to-one") %>% # join to building info by name
  filter(building != "CHW_Base") %>% # remove duplicate CHW value - not sure what this is?
  group_by(type_new, NAME, date, meter) %>% # group by building
  summarize(kwh = sum(kwh, na.rm = T),
            sqft = mean(sqft, na.rm = T)) %>%
  mutate(type = type_new) %>%
  filter(!is.na(type_new)) %>%
  ungroup() %>%
  select(type, meter, date, NAME, kwh, sqft) %>%
  mutate_all(~ifelse(is.nan(.), NA, .)) %>% # convert NaN to NA
  mutate(kwh = ifelse(kwh == 0, NA, kwh)) %>% # restore NA values for missing date-building combos
  filter(!is.na(date)) # remove rows without value for date

# generate daily summary for main and weis meters
daily_agg <- kwh %>%
  left_join(key, by = "meter_origin") %>%
  right_join(buildings_agg, by = "NAME") %>%
  group_by(type_new, NAME, date, meter) %>% # group by building
  summarize(kwh = sum(total_kwh, na.rm = T), # sum kwh by building
            sqft = mean(sqft, na.rm = T)) %>% # preserve sqft as is
  mutate(type = type_new) %>% 
  ungroup() %>%
  mutate_all(~ifelse(is.nan(.), NA, .)) %>%
  select(type, meter, date, NAME, kwh, sqft)

# generate complete daily dataset
 daily_full <- rbind(daily_individual, daily_agg, daily_sub) %>%
  mutate(date = as.Date(date)) %>%
  arrange(type, meter, NAME, date) %>%
  select(type, meter, NAME, date, kwh, sqft) %>%
  left_join(occupants, by = "NAME") %>%
  left_join(data_cov, by = "NAME") %>%
  merge(temp, by = "date") %>% # for some reason the join by date is not fully working... duplicating rows
  filter(!is.na(ave_temp)) %>% # remove duplicate rows
  mutate(period = ifelse((date > ymd("2024-08-28") & date < ymd("2024-12-21")), "Fall Semester", 
                         ifelse((date > ymd("2025-01-07") & date < ymd("2025-05-18")), "Spring Semester",
                                ifelse(date > ymd("2024-12-21") & date < ymd("2025-01-07"), "Winter Break", "Summer Break")))) %>%
   select(type, meter, NAME, days_perc, sqft, occupants, period, date, kwh, ave_temp) %>%
   arrange(NAME)
   
```

## Data availability by type

```{r graph meter, warning = F, fig.height = 5, fig.height = 4}
buildings_sum$meter <- factor(buildings_sum$meter,
                              levels = c("Weis Meter", "Main Meter",
                                        "Submeter", "Individual"))

pal <- c("lightgrey","darkgrey","#CAB2D6","#6A3D9A")

buildings_graph <- buildings_sum %>%
  filter(!is.na(meter) & !type_new %in% c("Res Hall - U","Non-building","Production","Mixed"))

ggplot(buildings_graph,
       aes(x = reorder(type_new, sqft, FUN = "sum"), y = sqft/1000, fill = meter)) +
  geom_col(position = "stack") +
  scale_fill_manual(values = pal) +
  theme_bw() +
  labs(x = "", y = "Square Footage (1000)", fill = "Meter status") +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

ggplot(buildings_graph, 
       aes(x = reorder(type_new, sqft, FUN = "sum"), fill = meter)) +
  geom_bar(position = "stack") +
  scale_fill_manual(values = pal) +
  theme_bw() +
  labs(x = "", y = "Number of Buildings", fill = "Meter status") +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
```

## Practice boxplot

Here's an example of how to make a faceted boxplot

```{r box_practice}
daily_test <- daily_full %>%
  filter(NAME %in% c("Main Meter","Weis Meter")) %>% 
  mutate(kwh_sqft = (kwh/sqft*365)) # adjust daily values to per year equivalent

ggplot(daily_test,
       aes(x = type, y = kwh_sqft, fill = type)) +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin = 7.4, ymax = 14.3, color = "lightgray", alpha = 0.3) +
  geom_hline(yintercept = 10.3, linetype = "dashed", color = "white") +
  geom_boxplot() +
  coord_flip() +
  facet_wrap(. ~ period) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(x = "", y = "Electricity Intensity (kWh/sqft)", 
       title = "Figure 2. Electricity intensity by aggregate meter in FY25. ")
```

## Export for downstream analyses

```{r export, results = 'hide'}
# final data check
str(daily_full)
str(joined_full_aca)
str(joined_full)

today <- today("EST") %>%
  str_remove(" UTC") 

write.csv(daily_full, paste0("../output/kwh_daily_", today, ".csv"), row.names = F)
write.csv(joined_full_aca, paste0("../output/kwh_academic_", today, ".csv"), row.names = F)
write.csv(joined_full, paste0("../output/kwh_annual_", today, ".csv"), row.names = F)
```

