---
title: "Final model visualization"
author: "Maggie Douglas"
date: "2026-04-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prepare data

```{r load, message = FALSE}
library(tidyverse)
library(segmented) # for segmented regression 
library(lmtest) # has functions for robust standard errors
library(sandwich) # has functions for robust standard errors

library(parameters) # print results table
library(gt) # format results table

library(visreg) # visualize regression results
library(dotwhisker) # create plots of effects from regression
```

```{r wrangle}
daily <- read.csv("./output/kwh_daily_2026-03-30.csv")

daily_date <- daily %>%
  mutate(date = ymd(date), # convert date to date object
         month = month(date, label = TRUE), # extract month
         day = wday(date, label = TRUE)) # extract day of week

# filter to the main meter
daily_main <- filter(daily_date, NAME == "Main Meter") 

# create factor for period and put in order
daily_main$period <- factor(daily_main$period,
                            levels = c("Fall Semester","Winter Break",
                                       "Spring Semester","Summer Break"))

# create factor for day, put in order, and store as a regular factor
daily_main$day <- factor(daily_main$day,
                         levels = c("Mon","Tue","Wed","Thu","Fri","Sat","Sun"),
                         ordered = FALSE)

str(daily_main)
```

## Fit the models

```{r mod}
# Original linear model
mod_full <- lm(kwh ~ day + period + ave_temp, data = daily_main)

# Segmented regression
mod_seg <- segmented(mod_full, seg.Z = ~ ave_temp, psi = 60)

# Quadratic model
mod_quad <- lm(kwh ~ day + period + ave_temp + I(ave_temp^2), data = daily_main)
```

## Visualize results

### Results table

When we are fitting a complex model, it is helpful to report the full results in a concise format. A table is a good way to do this. Below is an example for the segmented model; similar code should also work for standard or quadratic models. This function doesn't seem to have a built-in caption, so we will simply create a caption in our usual R Markdown text (remember that you can bold the table number by enclosing the text in two asterisks `**`)

**Table 1.** Statistical results of a segmented regression model for daily electricity use on the main meter. 

```{r res_tab}
results <- coeftest(mod_seg, vcov = vcovHAC(mod_seg)) 

print_html(model_parameters(results, summary = TRUE), digits = 0)
```

### Time of year

- This may throw a warning; it is okay to ignore.

```{r pred_period, warning = FALSE, fig.height = 4.5, fig.width = 5.5}
# Create a data frame with new values (must match model's predictor names)
period_data <- data.frame(ave_temp = 58, # set temp to the mean in the dataset
                       day = "Mon", # set day as Monday for consistency with other summaries
                       period = c("Fall Semester","Winter Break", # include all levels of period
                                  "Spring Semester","Summer Break"))

# Generate estimated means + confidence interval
conf_period <- as.data.frame(predict(mod_seg,  # store as a data frame, generate predictions from fitted model
                                     vcov = vcovHAC(mod_seg), # address the violations of assumptions
                                       newdata = period_data, # feed in the new data for which I want predictions
                                       interval = "confidence", level = 0.95)) %>% # set confidence interval to 95%
  mutate(period = c("Fall Semester","Winter Break",  # add period back into the data frame
                                  "Spring Semester","Summer Break"))

# Graph
ggplot(conf_period) +
    geom_bar(aes(x=period, y=fit), stat="identity", fill="skyblue", alpha=0.8) +
    geom_errorbar( aes(x=period, ymin=lwr, ymax=upr), width=0.4, colour="darkgray", alpha=0.9, size=1) +
  theme_bw() +
  labs(x = "", y = "Electricity use (kWh per day)", 
       title = "Figure 1. Main meter electricity use by time of year",
       subtitle = "Estimated for a Monday with an outdoor temperature of 58 degrees F")
```

### Day of the week

```{r pred_day, fig.height = 4.5, fig.width = 6, warning = FALSE}
# Create a data frame with new values (must match model's predictor names)
day_data <- data.frame(ave_temp = 58, 
                       day = c("Mon","Tue","Wed","Thu","Fri","Sat","Sun"),
                       period = "Fall Semester")

# Generate estimated means + confidence interval
conf_day <- as.data.frame(predict(mod_seg,  vcov = vcovHAC(mod_seg), 
                                       newdata = day_data, 
                                       interval = "confidence", level = 0.95)) %>%
  mutate(day = factor(c("Mon","Tue","Wed","Thu","Fri","Sat","Sun"),
                      levels = c("Mon","Tue","Wed","Thu","Fri","Sat","Sun")))

# Graph
ggplot(conf_day) +
    geom_bar(aes(x=day, y=fit), stat="identity", fill="skyblue", alpha=0.8) +
    geom_errorbar( aes(x=day, ymin=lwr, ymax=upr), width=0.4, colour="darkgray", alpha=0.9, size=1) +
  theme_bw() +
  labs(x = "", y = "Electricity use (kWh per day)", 
       title = "Figure 2. Main meter electricity use by day of the week",
       subtitle = "Estimated for a day in fall with an outdoor temperature of 58 degrees F")
```

### Temperature

#### Segmented model

The segmented model does not interact well with standard plotting functions, so we need to use a slightly different approach to its visualization. The code below uses R's base plotting function to visualize the fitted model.

```{r plot, fig.height = 5, fig.width = 6}
# plot original data with fit
plot(mod_seg, vcov = vcovHAC(mod_seg), 
     col = 'black', res = TRUE, conf.level = .95, shade = T, 
     res.col = adjustcolor("steelblue", alpha.f = 0.7), # set color of data + transparency
     xlab = "Temperature (degrees F)", ylab = "Electricity use (kWh per day)") # label axes
title("Figure 3. Main meter electricity use by temperature", # how to set the title in base R plotting function
      adj = 0, cex.main = 1.2, font.main = 1) 
```

#### Quadratic or linear model

For a linear model with or without a quadratic term, we can use a helpful function called `visreg` to visualize the results. Helpfully, by setting `gg = TRUE` the graph will be generated using ggplot, and then we can use all our usual functions to adjust its formatting. 

```{r visreg, warning = FALSE, fig.height = 4, fig.width = 6}
visreg(mod_quad, xvar = "ave_temp", # feed in model, specify temp as the variable
      cond=list(day = "Mon", period = "Fall Semester"), # set levels of other variables
       gg = TRUE) + # specify to use ggplot
  theme_bw() +
  labs(x = "Temperature (degrees F)", y = "Electricity use (kWh per day)",
       title = "Figure 3. Main meter daily electricity use by temperature")
```

### Comparing model estimates

Sometimes it's helpful to compare the estimates derived from different models to contextualize and interpret our results. Doing so helps us assess the sensitivity of our results to the modeling choices. The `dwplot` function allows us to easily compare the coefficients from a list of fitted models. 

```{r est, warning = FALSE}
dwplot(list(`Standard LM` = mod_full, # code on the left is the name of the model; right is the model object
            Segmented = mod_seg, 
            Quadratic = mod_quad)) + 
  theme_bw() +
  labs(x = "Diff. in daily kWh (compared to Monday in Fall)")
```
