---
title: "Main Meter - Simple linear regression"
author: "Maggie Douglas"
date: "2026-03-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question

What is the relationship between electricity use and outdoor temperature for buildings on the main meter?

## Load

```{r load}
library(tidyverse)
library(segmented) # for segmented regression 
library(tseries) # for time series, including autocorrelation

daily <- read.csv("./output/kwh_daily_2026-03-04.csv", strip.white = TRUE)
key <- read.csv("./keys/temp_by_date.csv")

str(daily)
str(key)
```

## Wrangle

```{r wrangle}
daily_main <- daily %>%
  filter(NAME == "Main Meter") %>%
  mutate(date = ymd(date),
         month = month(date, label = TRUE),
         day = wday(date, label = TRUE),
         day_type = ifelse(day %in% c("Sat","Sun"), "Weekend","Weekday")) 
```

## Visualize the relationship

```{r vis}
# basic relationship
ggplot(daily_main, aes(x = ave_temp, y = kwh)) +
  geom_point() +
  theme_bw() 

# what happens if we try to model this relationship with a straight line?
ggplot(daily_main, aes(x = ave_temp, y = kwh)) +
  geom_point() +
  theme_bw() +
  geom_smooth(method = "lm", formula = 'y ~ x')

# where is the break point?
ggplot(daily_main, aes(x = ave_temp, y = kwh)) +
  geom_point() +
  theme_bw() +
  geom_vline(xintercept = 58, col = "blue", linetype = "dashed")
```

## Fit a bad model (you wouldn't do this, but just to see what happens...)

```{r mod_bad}
mod_bad <- lm(kwh ~ ave_temp, data = daily_main)
```

### Check the fit

```{r check_bad}
par(mfrow = c(1, 2)) # This code put two plots in the same window
hist(mod_bad$residuals) # Histogram of residuals
plot(mod_bad, which = 2) # Quantile plot
plot(mod_bad, which = 1) # Residuals vs. fits
```

### Examine results

```{r results_bad}
summary(mod_bad) # Examine model terms + outcomes
```

## Fit a reasonable model

Let's focus here on temperatures over 58 degrees, where the relationship appears linear.

```{r mod}
# subset data > 58 degrees
daily_cool <- filter(daily_main, ave_temp > 58)

# fit the model for temps above 58
mod_cool <- lm(kwh ~ ave_temp, data = daily_cool)
```

### Check the fit

```{r check}
par(mfrow = c(1, 2)) # This code put two plots in the same window
hist(mod_cool$residuals) # Histogram of residuals
plot(mod_cool, which = 2) # Quantile plot
plot(mod_cool, which = 1) # Residuals vs. fits
```

### Examine results

```{r results}
summary(mod_cool) # Examine model terms + outcomes
```

### Plot results

```{r plot}
ggplot(daily_cool, aes(x = ave_temp, y = kwh)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = 'lm', formula = 'y ~ x') +
  theme_bw() +
  labs(x = "Temperature (degrees F)", y = "Electricity use (kWh)")
```

## What about a segmented regression?

This is a technique that essentially allows us to fit two linear models to our data, with a break point identified by the analysis. 

```{r mod_seg}
mod_full <- lm(kwh ~ ave_temp, data = daily_main)
mod_seg <- segmented(mod_full, seg.Z = ~ ave_temp, psi = 58)
```

### Check the fit

```{r check_seg}
par(mfrow = c(1, 2)) # This code put two plots in the same window
hist(mod_seg$residuals) # Histogram of residuals
plot(mod_seg$residuals) # Residuals vs. fits
```

### Examine results

```{r results_seg}
summary(mod_seg) # Examine model terms + outcomes
```

### Plot results

```{r plot_seg}
# plot original data with fit
plot(mod_seg, col = 'black', res = TRUE, conf.level = .95, shade = T, 
     res.col = adjustcolor("steelblue", alpha.f = 0.7),
     xlab = "Temperature (deg F)", ylab = "Electricity use (kWh)")
```
