---
title: 'PS 5: Prelim Academic Results'
author: 'Chloe Ross and Sophia Evans'
date: "2026-02-25"
output: html_document
---

## Data preparation

### Libraries

```{r load, results = "hide", message = F}
library(tidyverse)
library(DT)

```

### Annual electricty data

```{r annual, results = "hide"}
kwh_annual <- read.csv("./output/kwh_annual_2026-03-04.csv")
str(kwh_annual)

annual_academic <- kwh_annual %>%
  filter(type == "Academic") %>% 
  mutate(MTCO2e = round(kwh_corr*0.30082405, digits = 2), 
         Cost = round(kwh_corr*0.08138507,digits=2),
         kwh_sqft = round(kwh_corr/sqft, digits=2))%>%
  arrange(desc = TRUE) %>%
  select(meter, NAME, days_perc, kwh, kwh_corr, sqft, Cost, MTCO2e, kwh_sqft)
str(annual_academic)

```

### Daily electricity data

```{r daily, results = "hide"}
kwh_daily <- read.csv("./output/kwh_daily_2026-03-04.csv")
str(kwh_daily)
```

```{r wrangle, results = "hide"}
academic_kwh_daily <- kwh_daily %>% 
  filter(type == "Academic") %>% 
  mutate(date = ymd(date), 
         month = month(date, label = TRUE), 
         day = wday(date, label = TRUE)) %>%
  mutate(kwh_sqft = kwh/sqft*365)

str(academic_kwh_daily)

```

## Building type summary

### Descriptive table

```{r table}
colnames(annual_academic) <- c("Meter", "Building", "Percent of Days", "kWh per year", "kWh per year corrected", "sqft", "Cost ($)", "GHG emissions (MTCO2e)", "kWh per sqft")
datatable(annual_academic, filter = 'top', rownames = FALSE, caption = "Academic Building Summary Table")

```

### Electricity intensity

```{r intensity, warning = F}
academic_kwh_daily90 <- academic_kwh_daily %>%
  filter(days_perc >= 90)

ggplot(academic_kwh_daily90, aes(x = NAME, y = kwh_sqft, fill = NAME)) + 
         annotate("rect", xmin = -Inf, xmax = Inf, ymin = 7.4, ymax = 14.3, color = "lightgray", alpha = 0.3) + 
         geom_hline(yintercept = 10.3, linetype = "dashed", color = "white") + 
  geom_boxplot() + 
  coord_flip() + 
  theme_bw() + 
  theme(legend.position = "none") + 
  labs(x = "", y = "kWh per sqft per year", title = "Electricity Intensity by Academic Building") + 
  ylim(0, 30)
```

### Electricity use over the year

```{r use, warning = F}
ggplot(academic_kwh_daily90, aes(x = month, y = kwh, fill = NAME)) + 
  geom_col(position = "stack") + 
  theme_bw() + 
  labs(x = "", y = "Electricity use (kWh)", title = "Electricity use in Academic Buildings")

```

## Partner contributions

Sophia worked on annual kWh while I worked on daily kWh. We both worked on the table and the graphs (building type summary) together. 

Strange issues that we found is the very high outlier for Montgomery House which is 825.32521 kWh/sqft. The outliers are a problem for all submeter buildings. Because of the outliers we limited the y-axis to visualize the data. The data going missing at certain times makes the stacked bar graph inaccurate during certain months. 








