---
title: "PS05_Prelim_results"
author: 'Liam and Charlotte'
date: "2026-03-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Preparation    

### Conversion Factors  

```{r Conversion}
dol_per_kWh <- 0.081385
MTCO2_per_kWh <- 0.00030082405
```

### Libraries  

```{r libraries, results = "hide", message = F}
library(tidyverse) 
library(DT)
```  

### Annual Electricity Data   

```{r annual, results = "hide", message = F}
annual_kwh.df<-read.csv('./output/kwh_annual_2026-03-04.csv', strip.white=TRUE) #load data

str(annual_kwh.df) #check data

med_res_annual_kwh<-annual_kwh.df %>% #wrangle for building group annually
  filter(type=="Res Hall - M")



transform_med<-med_res_annual_kwh%>%
  mutate(
    cost=round(kwh_corr*dol_per_kWh, digits=0), 
    GHG_emis=round(kwh_corr*MTCO2_per_kWh, digits=0),
    kWh_per_sqft=round(kwh_corr/sqft, digits=1),
    kWh_per_person= round(kwh_corr/occupants, digits=0)
    )
  
  building_results<-transform_med %>%
    select(NAME, kwh_corr, kWh_per_sqft, cost, GHG_emis, kWh_per_person, meter, days_perc) %>%
  arrange(desc(kwh_corr))

```  
### Daily Electricity Data  

```{r daily, results = "hide", message = F}
med_res_daily_kwh <- read.csv("./output/kwh_daily_2026-03-04.csv")
str(med_res_daily_kwh)
med_res_daily_kwh <- med_res_daily_kwh %>%
  filter(type == "Res Hall - M") %>%
  mutate(date = ymd(date),
         month = month(date, label= TRUE),
         day = wday(date, label = TRUE)) %>%
  mutate(kwh_sqft_year = kwh/sqft*365, kwh_person_year = kwh/occupants*365) %>%
  arrange(desc(kwh))
```  



## Building Type Summary  
  
### Descriptive Table  

```{r table}

datatable(building_results,
          filter='top',
          rownames=FALSE,
          colnames = c("Building", "kWh", "kWh per sqft","Cost-$", "CO2e - MT", "kWh per person", "Meter", "Days of data (%)"),
          caption='Table 1. Annual Electricity Use Indicators by Medium Residence Hall in Fiscal Year 2025')
```  

### Electrcity Use Over the Year  '

```{r electricity use over the year, warning = F}
ggplot() +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin = 7.4, ymax = 14.3, color = "lightgray", alpha = 0.3) +
  geom_hline(yintercept = 10.3, linetype = "dashed", color = "white") +
  geom_boxplot(data = med_res_daily_kwh, aes(x = reorder(NAME, kwh_sqft_year, FUN = median), y = kwh_sqft_year, fill = NAME), show.legend = FALSE) +
  coord_flip() +
  labs(
    x = "",
    y = "Electricity Intensity (kWh/sqft/year)",
    title = "Electricity Intensity in Medium Residential Halls \nin the Fiscal Year 2025",
    subtitle = "Buildings ordered by median energy intensity") +
  theme_bw()

```  

### Electricity Intensity   

```{r intensity, warning = F}
ggplot(med_res_daily_kwh , aes(x=month, y=kwh, fill=NAME))+
  geom_bar(position="stack", stat="identity")+
  theme_bw()+
  labs(x="",
       y="Electricity Use (kWh)",
       fill= "Building",
       title= "Monthly Electricity Use by Medium Residence Hall in Fiscal Year 2025")
```

## Partner Contributions  


Charlotte prepared the annual electricity data for this problem set while Liam prepared the daily electricity data. For the building type summary, Charlotte adapted a table she had made last lab to create the summarizing table for annual electricity use indicators by building, Liam created the ranked box plot of daily electricity intensity by building, and Charlotte created the stacked bar graph to show electricity use by month. The data seems to have no issues  or problems, all the data is individually metered and clear. There is no missing data. There is occupancy data for all buildings and 100% of the days in the 2025 Fiscal Year were measured.  











