---
title: "PS 5: Preliminary results"
author: "Claire Raesly and Liv Kacar"
output: html_document
date: "2026-03-04"
---

## Data preparation

### Libraries

```{r load, results = "hide", message = F}

library(tidyverse)
library(DT)
library(paletteer)

```

### Annual electricity data

```{r annual, results = "hide", message = F}

annual <- read.csv("./output/kwh_annual_2026-03-04.csv")

str(annual)

cost_conversion <- 0.08138507
ghg_conversion <- 0.30082405 / 1000

annual_large_res <- annual %>%
  filter(type == "Res Hall - L") %>%
  mutate("Cost - $" = round(kwh_corr * cost_conversion, digits = 2), 
         "CO2e - MT" = round(kwh_corr * ghg_conversion, digits = 2), 
         "kWh per sqft" = round(kwh_corr/sqft, digits = 2),
         "kWh per person" = round(kwh_corr/occupants, digits = 2), 
         "Days of data (%)" = round(days_perc, digits = 2), 
         "kWh" = round(kwh, digits = 2),
         kWh_corr = round(kwh_corr, digits = 2), 
         "Meter" = meter, 
         "Building" = NAME) %>%
  arrange(desc(kwh_corr)) %>%
  select("Meter", "Building", "Days of data (%)", "kWh", "kWh_corr", "sqft", "kWh per sqft", "kWh per person", "Cost - $", "CO2e - MT")

str(annual_large_res)
summary(annual_large_res)

```

### Daily electricity data

```{r daily, results = "hide", message = F}

daily <- read.csv("./output/kwh_daily_2026-03-04.csv")

str(daily)

daily_large_res <- daily %>%
  filter(type == "Res Hall - L") %>%
  mutate(date = ymd(date),
         month = month(date, label = TRUE),
         day = wday(date, label = TRUE)) %>%
  mutate(kwh_tot_yr = kwh*365,
         kwh_sqft = kwh_tot_yr/sqft,
         kwh_person = kwh_tot_yr/occupants)

str(daily_large_res)
summary(daily_large_res)

```

## Building type summary

### Descriptive table

```{r table}

datatable(annual_large_res, filter = "top", rownames = FALSE, caption = "Large Residence Halls Summary")

```

### Electricity use over the year

```{r stackedbar, warning = F}

ggplot(daily_large_res, aes(x= month, y= kwh/(10^4), fill = NAME)) +
  geom_col(position = "stack") +
  theme_bw() +
  scale_fill_paletteer_d("ggthemes::Tableau_20")+
  labs(x = "",
       y = "Electricity use (millions of kWh)",
       fill = "Building",
       title = "Electricity Use by Building by Month", 
       subtitle = "Fiscal Year 2025")

```

```{r stackedbar_alt, warning = F, fig.height = 7, fig.width = 10}

ggplot(daily_large_res, aes(x= month, y= kwh/(10^4), fill = NAME)) +
  geom_col(position = "stack") +
  facet_wrap(. ~ NAME) +
  scale_fill_paletteer_d("ggthemes::Tableau_20")+
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(legend.position = "none") + 
  labs(x = "",
       y = "Electricity use (millions of kWh)",
       title = "Electricity Use by Building by Month", 
       subtitle = "Fiscal Year 2025")

```

```{r stackedbar_alt2, warning = F, fig.height = 7, fig.width = 10}

ggplot(daily_large_res, aes(x= month, y= kwh/(10^4), fill = NAME)) +
  geom_col(position = "stack") +
  facet_wrap(. ~ NAME, scales = "free_y") +
  scale_fill_paletteer_d("ggthemes::Tableau_20")+
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(legend.position = "none") + 
  labs(x = "",
       y = "Electricity use (millions of kWh)",
       title = "Electricity Use by Building by Month", 
       subtitle = "Fiscal Year 2025")

```

### Electricity intensity

```{r boxplot, warning = F}

ggplot(daily_large_res, aes(fct_reorder(NAME, kwh_sqft, median), y = kwh_sqft,
                            fill = NAME)) +
  annotate("rect", xmin=-Inf, xmax=Inf, ymin=8.2, ymax=36.1, 
           color="lightgrey", alpha= 0.3) +
  geom_hline(yintercept=13.6, linetype="dashed", color="white") +
  geom_boxplot() +
  ylim(0,50) +
  theme_bw() +
  theme(legend.position = "none") +
  scale_fill_paletteer_d("ggthemes::Tableau_20") +
  labs(x = "",
       y = "Electricity Intensity (kWh/sqft/year)",
       title = "Electricity Intensity by Square Foot (kWh/sqft/year)") +
  coord_flip()

```

## Partner contributions

**Claire**:

- Worked on the annual electricity data
- Made the summary datatable
- Wrote the partner contributions

**Liv**:

- Worked on the daily electricity data
- Made the graphs

**Data Issues**:

There are periods of time when a building meter goes offline, leading to a lot of NA values, then a very high value when the building comes back online. For now, we are adjusting the x-axis limit of the graph to better show the difference between buildings. Note that this means we have removed a lot of the outliers. Other than that, there were no issues.
