---
title: "Presentation_Community"
author: 'Dinela & Emma'
date: "2026-03-04"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

## Data Preparation 
### Libraries
```{r libraries, results = "hide", message = F}
library(tidyverse)
library(DT)
library(ggplot2)
kwh_yearly<-read.csv("./output/kwh_annual_2026-03-04.csv")
dailyfix<-read.csv("./output/kwh_daily_2026-03-04.csv")
```

### Annual Electricity Data
```{r annual_data, results = "hide", message = F}
community_yearly <- kwh_yearly %>%
  filter(type=="Community")%>%
  mutate(dollar_cost=round(kwh_corr*0.08138507,digits=2),
         ghg_emissions=round(kwh_corr*(0.30082405/1000),digits=4),
         kwh_per_sqft=round(kwh_corr/sqft, digits=4),
         kwh_per_person=round(kwh_corr/occupants, digits=4))%>%
  arrange(desc(kwh_corr))%>%
  select(-type)

str(community_yearly)
```

### Daily Electricity Data
```{r daily_data, results = "hide", message = F}
comm_dailyfix <- filter(dailyfix, type == "Community")%>%
  filter(days_perc>=95)%>%
  mutate(date = ymd(date), 
         month = lubridate::month(date, label = TRUE),
         day = lubridate::wday(date, label = TRUE)) %>%
  mutate(kwh_sqft = kwh*100/sqft)

str(comm_dailyfix)
```
  
  
## Building Type Summary
### Descriptive Table
```{r summary_tables}
datatable(community_yearly, caption="Community Building Annual Electricity Use Indicators")
```

### Electricity Use Over Year
```{r electricy/building/month, warning = F}            
ggplot(comm_dailyfix, aes(x = month, y = kwh, fill = NAME)) +
  geom_bar(position = "stack", stat = "identity") +
  scale_fill_brewer(palette = "RdYlBu")+
  theme_bw()+
  labs(x = "",
       y = "[Electricity use (kWh)]", 
       title = "Total Electricity Use by Building and Month", 
       fill = "Building")
```

### Electricity Intensity
```{r electricity_intensity, warning = F}
ggplot(comm_dailyfix, aes(x = reorder(NAME, kwh/sqft, FUN = median), y=kwh_sqft, fill=type))+
        geom_boxplot(show.legend=FALSE)+
  annotate("rect",xmin=-Inf, xmax=Inf, ymin=7.4, ymax=14.3, color="lightgray",alpha=0.3)+
    geom_hline(yintercept=10.3,linetype="dashed", color="white")+
    coord_flip()+
    theme_bw()+
    labs(x="",y="[Electricity Intensity (kWh/sqft/year)]",title="Electricity Intensity by Building",subtitle= "Ordered by Greatest Intensity")
```

## Partner Contributions  
  
  I, Dinela D., contributed to the code by transforming the annual electricity data and creating a table summarizing annual electricity use. 
    
  I, Emma K., worked on the daily subset of our data. I contributed to creating the stacked bar graph with total electricity use by building and month.
  
