---
title: "PS 5: Preliminary results"
author: "Claire Raesly and Liv Kacar"
output: html_document
date: "2026-03-04"
---

## Data preparation

### Libraries

```{r load, results = "hide", message = F}

library(tidyverse)
library(DT)
library(paletteer)

is_outlier <- function(x) {
  return(x < quantile(x, 0.25, na.rm = T) - 1.5 * IQR(x, na.rm = T) | x > quantile(x, 0.75, na.rm = T) + 1.5 * IQR(x, na.rm = T))
}

```

### Annual electricity data

```{r annual, results = "hide", message = F}

annual <- read.csv("./output/kwh_annual_2026-03-12.csv")

str(annual)

cost_conversion <- 0.08138507
ghg_conversion <- 0.30082405 / 1000

annual_large_res <- annual %>%
  filter(type == "Res Hall - L") %>%
  mutate("Cost - $" = round(kwh_corr * cost_conversion, digits = 0), 
         "CO2e - MT" = round(kwh_corr * ghg_conversion, digits = 0), 
         "kWh per sqft" = round(kwh_corr/sqft, digits = 1),
         "kWh per person" = round(kwh_corr/occupants, digits = 0), 
         "Days of data (%)" = round(days_perc, digits = 0), 
         "kWh" = round(kwh, digits = 0),
         "Corrected kWh" = round(kwh_corr, digits = 0), 
         "Meter" = meter, 
         "Building" = NAME) %>%
  arrange(desc(kwh_corr)) %>%
  select("Meter", "Building", "Days of data (%)", "kWh", "Corrected kWh", "sqft", "kWh per sqft", "kWh per person", "Cost - $", "CO2e - MT")

str(annual_large_res)
summary(annual_large_res)

```

### Daily electricity data

```{r daily, results = "hide", message = F}

daily <- read.csv("./output/kwh_daily_2026-03-12.csv")

str(daily)

daily_large_res <- daily %>%
  filter(type == "Res Hall - L") %>%
  mutate(date = ymd(date),
         month = month(date, label = TRUE),
         day = wday(date, label = TRUE)) %>%
  mutate(kwh_tot_yr = kwh*365,
         kwh_sqft = kwh_tot_yr/sqft,
         kwh_person = kwh_tot_yr/occupants) %>%
  group_by(NAME) %>%
  mutate(kwh_out = is_outlier(kwh)) %>%
  filter(!date %in% c(ymd("2024-09-02"),ymd("2024-09-06")))

outs <- filter(daily_large_res, kwh_out == TRUE)

str(daily_large_res)
summary(daily_large_res)

```

## Building type summary

### Descriptive table

```{r table}

datatable(annual_large_res, rownames = FALSE, 
          filter = "none",
          class = "compact",
          options = list(pageLength = 19, autoWidth = TRUE, dom = 't'),
          caption = "Table 1. Large Residence Halls summary of electricity use and associated cost and greenhouse gas emissions in fiscal year 2025. Corrected kWh reflects an estimate of annual electricity use including days with missing data. Occupancy reflects average student occupancy during the academic year.")

```

### Electricity use over the year

#### Option 1 - stacked bar

```{r stackedbar, warning = F}

ggplot(daily_large_res, aes(x= month, y= kwh/(10^3), fill = NAME)) +
  geom_col(position = "stack") +
  theme_bw() +
  scale_fill_paletteer_d("ggthemes::Tableau_20")+
  labs(x = "",
       y = "Electricity use (1000 kWh)",
       fill = "Building",
       title = "Electricity Use by Building by Month", 
       subtitle = "Fiscal Year 2025")

```

#### Option 2 - faceted, Y axis constant

```{r stackedbar_alt, warning = F, fig.height = 7, fig.width = 10}

ggplot(daily_large_res, aes(x= month, y= kwh/(10^3), fill = NAME)) +
  geom_col(position = "stack") +
  facet_wrap(. ~ NAME) +
  scale_fill_paletteer_d("ggthemes::Tableau_20")+
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(legend.position = "none") + 
  labs(x = "",
       y = "Electricity use (1000 kWh)",
       title = "Electricity Use by Building by Month", 
       subtitle = "Fiscal Year 2025")

```

#### Option 3 - faceted, Y axis varies

```{r stackedbar_alt2, warning = F, fig.height = 7, fig.width = 10}

ggplot(daily_large_res, aes(x= month, y= kwh/(10^3), fill = NAME)) +
  geom_col(position = "stack") +
  facet_wrap(. ~ NAME, scales = "free_y") +
  scale_fill_paletteer_d("ggthemes::Tableau_20")+
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(legend.position = "none") + 
  labs(x = "",
       y = "Electricity use (1000 kWh)",
       title = "Electricity Use by Building by Month", 
       subtitle = "Fiscal Year 2025")

```

### Electricity intensity

#### Option 1 - all days included

```{r boxplot, warning = F}

ggplot(daily_large_res, aes(fct_reorder(NAME, kwh_sqft, median), y = kwh_sqft,
                            fill = NAME)) +
  annotate("rect", xmin=-Inf, xmax=Inf, ymin=8.2, ymax=36.1, 
           color="lightgrey", alpha= 0.3) +
  geom_hline(yintercept=13.6, linetype="dashed", color="white") +
  geom_boxplot() +
 # ylim(0,50) +
  theme_bw() +
  theme(legend.position = "none") +
  scale_fill_paletteer_d("ggthemes::Tableau_20") +
  labs(x = "",
       y = "Electricity Intensity (kWh/sqft/year)",
       title = "Electricity Intensity by Square Foot (kWh/sqft/year)") +
  coord_flip()

```

#### Option 2 - only Academic Year

```{r boxplot2, warning = F}
daily_ay <- filter(daily_large_res, period %in% c("Spring","Fall"))

ggplot(daily_ay, aes(fct_reorder(NAME, kwh_sqft, median), y = kwh_sqft,
                            fill = NAME)) +
  annotate("rect", xmin=-Inf, xmax=Inf, ymin=8.2, ymax=36.1, 
           color="lightgrey", alpha= 0.3) +
  geom_hline(yintercept=13.6, linetype="dashed", color="white") +
  geom_boxplot() +
#  ylim(0,50) +
  theme_bw() +
  theme(legend.position = "none") +
  scale_fill_paletteer_d("ggthemes::Tableau_20") +
  labs(x = "",
       y = "Electricity Intensity (kWh/sqft/year)",
       title = "Electricity Intensity by Square Foot (kWh/sqft/year)") +
  coord_flip()

```

#### Option 3 - faceted by time of year

```{r boxplot3, warning = F, fig.height = 8, fig.width = 10}
daily_large_res$period <- factor(daily_large_res$period,
                                 levels = c("Fall","Spring","Winter","Summer"),
                                 labels = c("Fall Semester","Spring Semester",
                                            "Winter Break","Summer Break"))

ggplot(daily_large_res, aes(fct_reorder(NAME, kwh_sqft, median), y = kwh_sqft,
                            fill = NAME)) +
  annotate("rect", xmin=-Inf, xmax=Inf, ymin=8.2, ymax=36.1, 
           color="lightgrey", alpha= 0.3) +
  geom_hline(yintercept=13.6, linetype="dashed", color="white") +
  geom_boxplot() +
  facet_wrap(. ~ period) +
#  ylim(0,50) +
  theme_bw() +
  theme(legend.position = "none") +
  scale_fill_paletteer_d("ggthemes::Tableau_20") +
  labs(x = "",
       y = "Electricity Intensity (kWh/sqft/year)",
       title = "Electricity Intensity by Square Foot (kWh/sqft/year)") +
  coord_flip()

```


## Partner contributions

**Claire**:

- Worked on the annual electricity data
- Made the summary datatable
- Wrote the partner contributions

**Liv**:

- Worked on the daily electricity data
- Made the graphs

**Data Issues**:

There are periods of time when a building meter goes offline, leading to a lot of NA values, then a very high value when the building comes back online. For now, we are adjusting the x-axis limit of the graph to better show the difference between buildings. Note that this means we have removed a lot of the outliers. Other than that, there were no issues.
