---
title: "PS5: Preliminary Results"
author: Eva Hanlin and Amiya Marbles
date: "2026-03-04"
output: html_document
---

## Data Preparation

```{r library, results = "hide", message = F}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(readr)
library(RColorBrewer)
library(rmarkdown)
library(DT)
library(lubridate)
```

```{r data, results = "hide", message = F}
kwh_annual <- read_csv("./output/kwh_annual_2026-03-04.csv")
kwh_daily <- read_csv("./output/kwh_daily_2026-03-04.csv")
```

```{r check 1}
# str(kwh_annual)
# str(kwh_daily)
```

### Annual

Eva constructed the code for annual data elements including building category, transform, and organize.

```{r building category, results = "hide"}
annual_admin <- filter(kwh_annual, type == "Admin")
```

```{r conversion factors}
ghg_kg <- 0.30082405
cost <- 0.08138507
```

```{r transform, results = "hide"}
final_annual_admin <- annual_admin %>%
  mutate(GHG.MT = round(kwh_corr*ghg_kg)/1000,
         cost = round(kwh_corr*cost),
         kwh_sqft = round(kwh/sqft))
final_annual_admin
```

```{r organize, results = "hide"}
final_annual_admin %>%
  arrange(final_annual_admin, kwh_corr)
final_annual_admin$type <- NULL
```

```{r check, results = "hide"}
str(final_annual_admin)
# summary(final_annual_admin)
```

### Daily

Amiya conducted the analysis using code for the daily data of administrative buildings including filtering for administrative buildings and transforming the data to the necessary measurements.

```{r daily}
Admin_subset <- kwh_daily %>%
  filter(type == "Admin") %>%
  mutate(
    date = ymd(date),
    month = month(date, label = TRUE, abbr = FALSE),
    day_of_week = wday(date, label = TRUE, abbr = TRUE),
    kwh_sqft_yr = round(kwh*365) / sqft)
```

## Building Type Summary - Admin

Eva created the descriptive tables for the administrative buildings and created the stacked bar chart for electricity intensity by month. It is important to note for the stacked bar chart that there is data missing for all the buildings on the Weiss Meter, including Old West, for part of August, making Old West's total kWh for the month much lower than reality. Amiya created the ranked boxplot for annual total kwh by administrative building using the EIA 2018 report measurements of spread. 

```{r descriptive table}
datatable(final_annual_admin)
datatable(Admin_subset)
```

```{r electricity intensity boxplot, warning = F}
eia_median <- 10.1
eia_q25 <- 6.1    
eia_q75 <- 15.7   
  ggplot(Admin_subset,aes(
    x = reorder(NAME, kwh_sqft_yr, FUN = median, na.rm = TRUE), 
    y = kwh_sqft_yr,
    fill = NAME
  )) +
  annotate("rect", ymin = eia_q25, ymax = eia_q75, xmin = -Inf, xmax = Inf, 
           alpha = 0.2, fill = "gray50") +
  geom_hline(yintercept = eia_median, linetype = "dashed", color = "black", linewidth = 0.8) +
  geom_boxplot(alpha = 0.9, outlier.alpha = 0.5) +
  coord_flip() +
  facet_wrap(~ period) +
  labs(
    title = "Annualized Daily Electricity Intensity by Building",
    subtitle = "Dashed line = EIA 2018 Median; Gray band = EIA 2018 IQR",
    x = NULL, 
    y = "Electricity Intensity (kWh/sqft/year)"
  ) +
  theme_bw() +
  theme(
    legend.position = "none",
    panel.spacing = unit(1, "lines")
  )
```

```{r electricity intensity stacked bar, warning = F}
ggplot(Admin_subset, aes(x = month, y = kwh, fill = NAME)) +
  geom_col() +
  theme_bw() +
  theme(legend.title = element_blank()) +
  labs(title = "Electricity Intensity by Month for Administrative Buildings",
       x = "Month",
       y = "Administrative Building Name/Address")
```
