---
title: "Academic Buildings"
author: "Sophia Evans and Chloe Ross"
date: "`r format(Sys.time(), '%m/%d/%y')`"
output: html_document
---

## Background

The buildings in the academic category primarily hold classrooms, laboratories, offices and study spaces. They are mostly used by students and faculty during work hours on weekdays. Many of the buildings are still used by a smaller number of people at nights and on weekends. The buildings in this category cover 517,099 sqft on campus which is about 27% of building space. These buildings very greatly in size. The largest building is the Waidner Spahr Library, which is 120,000 sqft, and the smallest is the Keck Archaeology lab, which is 1,511 sqft. This category also has significant variability in age, from 19 years to 199 years old.

This category contains 20 buildings; however, only half of these have individual data. Five academic buildings are metered individually, and 15 are on the main meter. Of the buildings on the main meter, only five buildings have usable daily submeter data. This presents a significant challenge to understanding the electricity use of academic buildings on campus since only half of the buildings have usable data. Additionally, these buildings only account for 45% of the square footage of the academic buildings. The following analysis is conducted only on the ten buildings with individual daily data. 

```{r load, include = FALSE}
library(tidyverse)
library(DT)
library(paletteer)
kwh_academic_year <- read.csv("./output/kwh_academic_2026-03-17.csv")
kwh_annual <- read.csv("./output/kwh_annual_2026-03-17.csv")
kwh_daily <- read.csv("./output/kwh_daily_2026-03-17.csv")
```

```{r check, include = FALSE}
str(kwh_annual)
str(kwh_daily)
str(kwh_academic_year)
```

```{r annual_academic, include = FALSE}
annual_academic <- kwh_annual %>%
  filter(type == "Academic") %>% 
  mutate(MTCO2e = round(kwh_corr*0.00030082405, digits = 1), 
         Cost = round(kwh_corr*0.08138507,digits= 0),
         kwh_sqft = round(kwh_corr/sqft, digits= 1),
         days_perc = round(days_perc, digits = 0),
         kwh = round(kwh, digits = 0),
         kwh_corr = round(kwh_corr, digits = 0)
         )%>%
  arrange(-kwh) %>%
  select(meter, NAME, days_perc, sqft, kwh, kwh_corr, kwh_sqft, Cost, MTCO2e)
str(annual_academic)
```

```{r ayear_academic, include = FALSE}
ayear_academic <- kwh_academic_year %>%
  filter(type == "Academic") %>% 
  mutate(MTCO2e = round((kwh_corr*0.00030082405), digits = 1), 
         Cost = round(kwh_corr*0.08138507,digits= 0),
         kwh_sqft = round(kwh_corr/sqft, digits= 1),
         days_perc = round(days_perc, digits = 0),
         kwh = round(kwh, digits = 0),
         kwh_corr = round(kwh_corr, digits = 0))%>%
  arrange(-kwh) %>%
  select(meter, NAME, days_perc, sqft, kwh, kwh_corr, kwh_sqft, Cost, MTCO2e)
str(ayear_academic)
```

```{r daily, include = FALSE}
academic_kwh_daily <- kwh_daily %>% 
  filter(type == "Academic") %>% 
  mutate(date = ymd(date), 
         month = month(date, label = TRUE), 
         day = wday(date, label = TRUE)) %>%
  mutate(kwh_sqft = kwh/sqft*365) %>%
  arrange(-kwh)
str(academic_kwh_daily)
```

## Electricity Use Summary

For the academic buildings the total annual electricity consumption was highly variable by building. During fiscal year 2025, the median annual corrected electricity consumption for academic buildings was 34,465 kWh with an interquartile range from 14,560 kWh to 118,119 kwh. These buildings also had significant variability in electricity cost and GHG emissions. The estimated cost of electricity per building in fiscal year 2025 ranged from $528 to $29,356. Similarly, the estimated GHG emissions ranged from 2 MT CO2e to 109 MT CO2e. This variability is expected considering the range of sizes that this category covers. The square footage ranged from 1,775 sqft to 120,000 sqft (Table 1). However, electricity intensity was less variable. The median annual electricity intensity was 3.7 kWh/sqft and the interquartile range was 3.6 to 4.7 kWh/sqft. 

```{r table1}
datatable(annual_academic, 
          filter = 'none', 
          colnames = c("Meter", "Building", "Days of data (%)", "Square footage", "kWh", "kWh (corrected)","kWh per sqft", "Cost ($)", "CO2e (MT)"),
          rownames = FALSE, 
          class = "compact",
          options = list(pagelength = 10, autoWidth = TRUE, dom = 't'),
          caption = "Table 1. Total electricity use, estimated financial cost, and estimated greenhouse gas emissions of academic buildings during fiscal year 2025")
```

The academic buildings are used differently throughout the year. During breaks, the academic buildings see less use while most of the student body is away from campus and classes are not in session. Thus, it is useful to examine the data excluding breaks to better characterize electricity use during the academic year. Table 2 shows electricity consumption by building during the academic year 2024-2025. Both electricity consumption and electricity intensity were lower during only the academic year and there is still significant variability between buildings. 

```{r table2}
datatable(ayear_academic, 
          filter = 'none', 
          colnames = c("Meter", "Building", "Days of data (%)", "Square footage", "kWh", "kWh (corrected)","kWh per sqft", "Cost ($)", "CO2e (MT)"),
          rownames = FALSE, 
          class = "compact",
          options = list(pagelength = 10, autoWidth = TRUE, dom = 't'),
          caption = "Table 2. Total electricity use, estimated financial cost, and estimated greenhouse gas emissions of academic buildings during academic year 2024/2025")
```

The largest electricity users among Dickinson's academic buildings with electricity data are
Waidner-Spahr Library, Althouse, Bosler, and the Stern Center, which consistently showed the
highest total monthly electricity consumption, particularly during the academic year (Figure 1). Buildings such as 46 S. West St., 162–164 Dickinson Ave. and 57 S. College St. were the smallest electricity users overall, with consistently low consumption across all months (Figure 1). There was substantial variability in electricity use both across buildings and across seasons. Larger, high-use buildings maintained relatively high consumption year-round but often showed slight decreases during summer or winter breaks. Some buildings had a dip in the month of August that was likely from missing data for submetered buildings (Figure 1). Other buildings, such as Montgomery House and Waidner-Spahr Library (which are both on submeters), exhibited extreme variability. For Montgomery House there was a sharp spike in December that was inconsistent with the rest of the year while for the Waidner-Spahr Library there was only data for half of the year (Figure 1). These buildings with missing or inconsistent electricity data could be reflecting an issue with the submeter data.

```{r figure1, warning = FALSE}
ggplot(academic_kwh_daily, aes(x = month, y = kwh/(10^3), fill = NAME)) + 
  geom_col(position = "stack") + 
  facet_wrap(. ~ NAME) + 
  scale_fill_paletteer_d("ggthemes::Tableau_20") + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  theme(legend.position = "none") + 
  labs(x = "",
       y = "Electricity use (1000 kWh)",
      title = "Figure 1. Monthly patterns in electricity use by building for academic buildings \nin FY25")
```

When compared to the EIA 2018 Commercial Buildings Energy Consumption Survey, Dickinson academic buildings with usable electricity data fell below the EIA Survey’s national median of typical values for U.S. colleges and universities (10.3 kWh/sqft) and within or below the interquartile range (7.4 kWh/sqft - 14.3 kWh/sqft). This indicates generally strong performance with the exception of Green Valley Sanctuary and Carlisle Theatre, which had the highest electricity intensities across multiple seasons (Figure 2), often approaching or exceeding the 75th percentile (14.3 kWh/sqft). The buildings with the highest electricity intensity were Green Valley Sanctuary and Carlisle Theatre, which consistently fell at or above the median and the 75th percentile of typical values across multiple seasons (Figure 2). These patterns were particularly acute during the academic year (fall and spring semesters). Montgomery House and 57 S. College St. were among the lowest-intensity buildings across all seasons, suggesting relatively low electricity use per square foot (Figure 2). Other academic buildings (Althouse, Bosler Hall, Stern Center, and Waidner-Spahr Library) fell within the typical range (7.4 kWh/sqft - 14.3 kWh/sqft) of electricity intensity, and were generally below the national median (Figure 2). There is notable variability in electricity intensity across both buildings and seasons. Some buildings, such as the Carlisle Theatre, showed a wide spread of values, particularly during the summer (Figure 2). Others, like Green Valley Sanctuary, maintained consistently high intensity across all periods (Figure 2). Seasonal variation was also evident, with many buildings showing slightly lower intensity during summer and winter breaks, although this pattern was not uniform across all buildings. 

```{r figure2, warning = FALSE}
ggplot(academic_kwh_daily, aes(fct_reorder(NAME, kwh_sqft, median), y = kwh_sqft,
                            fill = NAME)) + 
         annotate("rect", xmin = -Inf, xmax = Inf, ymin = 7.4, ymax = 14.3, color = "lightgray", alpha = 0.3) + 
         geom_hline(yintercept = 10.3, linetype = "dashed", color = "white") + 
  geom_boxplot() +
  facet_wrap(. ~ period) + 
  coord_flip() + 
  theme_bw() + 
  theme(legend.position = "none") + 
  scale_fill_paletteer_d("ggthemes::Tableau_20") +
  labs(x = "", 
       y = "Electricity Intensity (kWh/sqft/year)", 
       title = "Figure 2. Electricity intensity by building for academic buildings \nin FY25",
      caption = "White dotted line and gray shaded area show median and 25th to 75th percentile \nof typical values in U.S. College and University buildings (EIA 2022)") +
  ylim(0,30) # Montogomery House as an outlier of 800
```

## Recommendations 

A limitation of this study is the inconsistency in submeter data availability across academic buildings. As shown in Tables 1 and 2, several buildings did not have complete daily records due to meters going offline, which can result in unrecorded electricity use being aggregated into later days when the meter goes back online. To improve data quality, Dickinson should prioritize upgrading or replacing submeters in buildings with significant data gaps, particularly Waidner-Spahr Library, which is the largest building in this category (120,000 sqft) and had only 36% of days with electricity data despite being heavily used throughout the academic year (Table 1). Montgomery House is another building we recommend looking into because it exhibits an extreme outlier in electricity intensity (approximately 800 kWh/sqft/year) and unusually high electricity use just in the month of December (Figure 1). This pattern may reflect submeter problems or delayed reporting of electricity use due to going offline, but it could also indicate underlying issues with electricity use.

When compared to the U.S. Energy Information Administration’s 2018 Commercial Buildings Energy Consumption Survey, Dickinson’s academic buildings that have usable electricity data perform well, with electricity intensities below the national median during both academic semesters and break periods. However, exceptions such as Carlisle Theatre and Green Valley Sanctuary meet or exceed the EIA median (Figure 2), suggesting potential problems or different electricity demands that should be further examined. Dickinson’s academic buildings are functioning well and are showing a generally strong performance compared to the college and university section in the EIA building survey, but variation between buildings suggests there are still opportunities to improve electricity use. Althouse, Bosler, Stern Center, and Waidner-Spahr Library consistently show higher monthly electricity use than others (Figure 1) but these buildings are also the largest buildings in the academic category (Table 1), which contributes to more electricity consumption. It is important to note that these buildings also fall toward the middle of the distribution for electricity intensity when compared to other academic buildings and well below the EIA ranges for typical college buildings (Figure 2). Even though these buildings don’t have the highest electricity intensity, improvements to these buildings most likely have the largest pay-off in terms of electricity savings due to their size and everyday use by students and faculty. Buildings with consistently low electricity intensity relative to other buildings may serve as useful models, and analyzing their operational practices could help inform electricity-saving strategies
across campus.


## Sources

Energy Information Administration (EIA). (2022). 2018 Commercial Buildings Energy Consumption Survey (CBECS). https://www.eia.gov/consumption/commercial/

Leary, N. (2025). Dickinson College Greenhouse Gas Inventory 2008-2023. Center for Sustainability Education.

## Partner Contributions

Sophia wrote the background, generated the tables and wrote the associated text. Chloe wrote the recommendations, generated the figures and wrote the associated text. Both partners contributed to revisions.
