---
title: "Other Buildings"
author: "Ben Connors & Kyrie Sanchez"
date: "2026-03-18"
output: html_document
---

## Background

The 'other' category is the accumulation of buildings that do not align with any one group. There are 17 buildings in total, which make up roughly 13% of buildings on campus. The total square feet across every building category is 1,946,878 square feet; buildings listed as other have a combined total of 250,130 square feet which is also 13% of building space.  

The buildings in the other category are incredibly variable. The largest building is the Kline, which contributes 20% of the category's square footage, whereas the smallest, 155 W. High St., Ground Floor, contributes less than 1%. There is also a lack of consistency of the age of the buildings, when they were last renovated, and what heating/cooling system they use. 

Some of the buildings may host residents but are not exclusively residential – such as Kade House and King House. Some buildings represent multiple structures whose electricity consumption is tracked by a single electric meter. For example: Biddle Press Box, Restrooms, and Durden ATC (Biddle) are on one meter. 5 N. Orange St., which is one building with two meters, is split with the student group focusing on administrative buildings. The half of the building represented by this category is used by ProjectSHARE and the Central Pennsylvania Youth Ballet (CPYB). As a result of the diversity within the category, it is hard to identify a main function or building use that contributes to electricity use. 

A majority (13/17) of the other buildings have complete individual data sets for electricity use in fiscal year 2025. Buildings that have significant missing data have been removed from our analysis. Kline, which has been disaggregated from the main meter, only has 35% of daily data. No electricity data is available for Boiling Springs Cottage and Boiling Springs Farm. 25 Pomfret Street and 50 Wilson Street are missing square footage data. 

```{r load, include = FALSE}
library(tidyverse)
library(DT)
library(paletteer)
annual_data <- read.csv("./output/kwh_annual_2026-03-17.csv")
academic_data <- read.csv("./output/kwh_academic_2026-03-17.csv")
daily <- read.csv("./output/kwh_daily_2026-03-17.csv")
```

```{r check, include = FALSE}
str(annual_data)
str(academic_data)
str(daily)
```

```{r subset_annual, include = FALSE}
annual_other <- annual_data %>%
   filter(type == "Other") %>%
  mutate(kwh_per_sqft = round(kwh_corr / sqft, digits = 1),
    cost_dol = round(kwh_corr * .08138507),
         CO2e_MT = round(.30082405*kwh_corr/1000, digits = 1),
         days_perc = round(days_perc),
         kwh = round(kwh),
         kwh_corr = round(kwh_corr),
         building = NAME)%>%
  arrange(desc(kwh_corr)) %>%
  select(meter, building, days_perc, sqft, kwh, kwh_corr, kwh_per_sqft, cost_dol, CO2e_MT) %>%
filter(building != "Kline Life Sport Complex")
```

```{r check_annual, include = FALSE}
str(annual_other)
```

```{r subset_academic, include = FALSE}
academic_other <- academic_data %>%
  filter(type == "Other") %>%
  mutate(kwh_per_sqft = round(kwh_corr / sqft, digits = 1),
    cost_dol = round(kwh_corr * .08138507),
         CO2e_MT = round(.30082405*kwh_corr/1000, digits = 1),
         days_perc = round(days_perc),
         kwh = round(kwh),
         kwh_corr = round(kwh_corr),
         building = NAME)%>%
  arrange(desc(kwh_corr)) %>%
  select(meter, building, days_perc, sqft, kwh, kwh_corr, kwh_per_sqft, cost_dol, CO2e_MT) %>%
  filter(building != "Kline Life Sport Complex")
```

```{r check_academic, include = FALSE}
str(academic_other)
```

```{r daily, include = FALSE}
other.daily<-daily%>%
  filter(type=="Other")%>%
  mutate(date=ymd(date),
        day=wday(date,label=TRUE),
        month=month(date,label=TRUE))%>%
  mutate(kwh.total=kwh*365,
    kwh.sqft.year=kwh.total/sqft)%>%
  filter(NAME !="Kline Life Sport Complex", NAME != "Boiling Springs Cottage", NAME != "Boiling Springs Farm")
# Kline removed due to the building only having data for 37% of days
# Boiling Springs Cottage and Farm removed due to only temperature data being available

str(other.daily)
```

## Electricity use summary

There is a lot of variation in square footage and usage patterns in the other building category, and thus, there was a large spread in electricity use for this category (Table 1). The median value across the entire fiscal year was 19,990 kWh (interquartile range 9,118 – 44,548 kWh). The median electricity intensity in this category was 6.5 kWh/sqft (interquartile range 4.5 – 9.6 kWh/sqft). 

The median electricity cost for a building in the other category was $1,226 per year (interquartile range $432 – $2,961 per year). However, the mean electricity cost for a building in this category was $3,399, which is significantly higher than the median. This is because the two highest individually metered users in the category, Biddle and ProjectSHARE, used significantly more electricity compared to the rest of the other buildings.

The median greenhouse gas emissions in the other category was 5 metric tons of CO2 equivalent (interquartile range 2 – 11 MTCO2e), with a mean of about 13 MTCO2e . While the median values better represent most of the buildings in the category, the mean value is still important for demonstrating how much influence the biggest electricity consumers have on this category.

Biddle and Project SHARE were the two biggest overall electricity users during the 2025 fiscal year, using 298,475 and 277,590 kWh respectively (Table 1). These values were much higher than King House, the third highest electricity consumer, which used 64,445 kWh throughout the year. However, King House was the most electricity intensive user, as when the data were normalized by square footage, the building used 23.5 kWh/sqft. Biddle was the second most electricity intensive building, and ProjectSHARE was the sixth most intensive, with values of 12.0 and 6.5 kWh/sqft respectively. 155 W High St Ground Floor and 368 W. North were the two smallest total electricity users over the fiscal year, using 1068 and 3151 kWh respectively. Wentzel was the building with the lowest electricity intensity, using only 0.9 kWh/sqft. There is no square footage data for 25 Pomfret St or 50 Wilson St, so these two buildings were not included in the normalization analysis.
	

```{r table_one}
datatable(annual_other, caption = "Table 1. Total electricity use, estimated financial cost, and estimated greenhouse gas emissions of other buildings during fiscal year 2025 ",
          rownames = FALSE,
          colnames = c("Meter", "Building", "Days of data (%)", "Square footage", "kWh", "kWh (corrected)", "kWh per sqft", "cost ($)", "CO2e (MT)"),
          filter = "none",
          class = "compact",
          options = list(pageLength = 13, autoWidth = TRUE, dom = 't'))
```

```{r table1_sum, include = FALSE}
summary(annual_other)
```

During the academic year, Biddle and ProjectSHARE were still by far the highest electricity users in the other category (Table 2). However, ProjectSHARE was the leading user, using 192,318 kWh, while Biddle used 190,898 kWh. King House was still the third highest, using 46,665 kWh. King House and Biddle were still the most electricity intensive buildings during the academic year, using 17.0 and 7.7 kWh/sqft respectively. ProjectSHARE was the fifth most electricity intensive building, using 4.5 kWh/sqft. During the academic year, 155 W High St Ground Floor and 368 W North were still the two lowest total electricity users, using 782 and 2206 kWh respectively. Wentzel was still the least electricity intensive building, using 0.7 kWh/sqft. 

```{r table_two}
datatable(academic_other, caption = "Table 2. Total electricity use, estimated financial cost, and estimated greenhouse gas emissions of other buildings during academic year 2024/2025",
          rownames = FALSE,
          colnames = c("Meter", "Building", "Days of data (%)", "Square footage", "kWh", "kWh (corrected)", "kWh per sqft", "cost ($)", "CO2e (MT)"),
          filter = "none",
          class = "compact",
          options = list(pageLength = 13, autoWidth = TRUE, dom = 't'))
```

```{r table2_sum, include = FALSE}
summary(academic_other)
```

There was a lot of variation in seasonal patterns among the other category buildings (Figure 1). Buildings such as Biddle and 362 W. North St displayed patterns of peak electricity use in the summer. Other buildings like King House and 25 Pomfret St displayed the opposite pattern, with electricity use peaking in the winter, and being the lowest in the summer months. Wentzel also had a unique pattern, peaking in the fall during the months of September-November. All of these buildings also displayed very different values for electricity usage due to the variation in square footage and usage within the building category.  

```{r figure_one, fig.width = 9.5, fig.height = 7}
ggplot(other.daily, aes(x=month, y=kwh/10^3, fill=NAME))+
  geom_col(stat = "identity", position = "stack")+
  facet_wrap(. ~ NAME) +
  scale_fill_paletteer_d("ggthemes::Tableau_20")+
  labs(title="Figure 1. Monthly patterns in electricity use in 'Other' buildings",
  subtitle = "Fiscal Year 2025",
  x="",
  y="Electricity use (1000 kWh)",
  fill="Building")+
  theme_bw()+
  theme(strip.text = element_text(size = 8)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  theme(legend.position = "none")
```

There was also a lot of variation in electricity intensity in the other category buildings, but the buildings displaying high or low electricity intensity were generally consistent throughout the year (Figure 2). Throughout the entire fiscal year, King House was consistently the most electricity intensive building, with a significantly higher median and interquartile range (IQR) compared to the rest of the buildings in this category. These values were also much higher than the median and IQR for similar buildings at other colleges and universities (EIA 2022). Biddle was consistent throughout the year, despite a slight peak in electricity intensity during summer break. This value is comparable to similar buildings at other colleges and universities. Most of the other buildings were below the comparison values, but in the summer, there were five buildings that were similar or higher: King House, Biddle, 344 W North St, 362 W North St, and Kade House. 49 W High St and 119 W High St saw noticeable peaks in electricity intensity during the spring semester and winter break.

``` {r figure_two, fig.width = 10, fig.height = 6.5}
other.box<-other.daily%>%
  filter(!is.na(sqft))
ggplot(other.box,aes(x=reorder(NAME,kwh.total/sqft,FUN=median),y=kwh.sqft.year,fill=NAME))+
    annotate("rect", xmin=-Inf, xmax=Inf, ymin=7.4, ymax=14.3, 
           color="lightgrey", alpha= 0.3) +
  geom_hline(yintercept=10.3, linetype="dashed", color="white") +
  geom_boxplot()+
  facet_wrap(. ~ period) +
  labs(title="Figure 2. Electricity Intensity by building for other buildings in FY25",
  subtitle="Only includes buildings with individual meters, and square footage data",
  x="",
  y="Electricity Intensity (kWh/sqft/year)",
  caption = "White dotted line and gray shaded area show median and 25th to 75th percentile of typical values in U.S. College and University buildings (EIA 2022)")+
  coord_flip()+
  theme_bw()+
  theme(legend.position="none")+
  scale_fill_paletteer_d("ggthemes::Tableau_20")

# median and IQR come from EIA (2022) table, annual kWh per square foot for Colleges/Universities
# https://www.eia.gov/consumption/commercial/data/2018/ce/pdf/c22.pdf
```

## Recommendations

Based on the total electricity consumption for this category, Biddle is the most intriguing building to explore further. What makes this building complex so interesting is its peak in total use during the summer months. There is likely not as much use for this building complex during the summer, so this could potentially be an easy way to cut down on a large amount of electricity. Project SHARE might also be interesting to investigate because of the large amount of electricity it consumes, but since the electricity intensity is not very high, and is lower compared to similar buildings at other colleges and universities, this building may be operating reasonably efficiently already. It is unlikely that we could find a single model for this building category given the variance in seasonal patterns, electricity intensity, and total electricity usage.

Another pattern that is interesting in the data is the consistently very high electricity intensity at King House. This is substantially higher than the other buildings in the category and even surpasses the comparison value from other colleges and universities. This building could likely have much more efficient electricity usage. Wentzel’s electricity use peaking in the fall is also an interesting pattern, and it could be valuable to look more into this building to find out why this is, and if the electricity use could be lowered during the peak months. However, since the total usage of this building is not very high, it will likely not be a priority of further analysis.

Since there is such pronounced variability within this building category, to improve our understanding of electricity usage patterns, it would be helpful to know what the heating and cooling systems are for each building. This would help establish building-specific expectations and identify anomalies. 

Due to a lack of similarities within this category, we cannot effectively model electricity use for the other buildings as a conglomerate. Instead, it is more valuable to investigate individual buildings of interest and report on how those single buildings can be improved. Project SHARE is using electricity in a predictable way given the building's layout and purpose despite being one of the largest electricity consumers. However, Biddle has unusually high use at all times where we would expect to see a decrease in electricity consumption in the summer, suggesting that there is room for improvement. Other individual buildings have not been observed as heavily as Project SHARE and Biddle. 

## Sources

Energy Information Administration (EIA). (2022). *2018 Commercial Buildings Energy Consumption Survey (CBECS).* [https://www.eia.gov/consumption/commercial/](https://www.eia.gov/consumption/commercial/)

Leary, N. (2025). *Dickinson College Greenhouse Gas Inventory 2008-2023.* Center for Sustainability Education.

## Partner contributions

Ben Connors performed data wrangling and created the tables, working with the annual and academic datasets. He also wrote the electricity use summary. Kyrie Sanchez worked performed data wrangling on the daily data and created the two figures. She also wrote the background. Both partners contributed to the recommendations and revisions. No AI tools were used for any part of this report.
