---
title: "Adminstrative Buildings"
author: "Amiya Marbles & Eva Hanlin"
date: "`r format(Sys.time(), '%m/%d/%y')`"
output: html_document
---

## Background

There are eleven total administrative buildings listed in this campus category, accounting for a combined **188,060 square feet** of space. However, since the electricity dataset only covers *six* of these buildings - five with individual meters and Old West on a submeter - we are actively analyzing **137,948 square feet**, which is about **73%** of the total administrative footprint. Almost all of these administration buildings are used on a typical 9-5PM schedule year-round. They cover a highly diverse range of campus activities, including executive functions in Old West, campus operations departments like Facilities Management, and program associated buildings like the ROTC offices. There is a massive spread in the physical size of these buildings, which may impact electricity usage. 5 North Orange Street - containing Facilities Management, Office of Conferences and Special Events, and the Print Center - is by far the largest at 75,000 square feet, followed by Old West at 30,208 square feet, while the smallest spaces analyzed are converted residential structures like the 1,440 square feet ROTC office. Additionally, the architecture spans over two centuries, meaning energy efficiency and insulation may vary depending on renovations to HVAC systems since original build, ranging from historic Old West built in 1803 to mid-century builds and recently renovated spaces like the ROTC offices at 554 and 556 West Louther Street.

```{r data_prep, include = FALSE}
library(dplyr)
library(readr)
library(RColorBrewer)
library(tidyverse)
library(DT)
library(paletteer)
library(lubridate) 
library(rmarkdown)
library(ggplot2)

kwh_annual_df <- read_csv("./output/kwh_annual_2026-03-17.csv")
kwh_daily <- read_csv("./output/kwh_daily_2026-03-17.csv")
kwh_academic <- read.csv("./output/kwh_academic_2026-03-17.csv")
# str(kwh_annual)
# str(kwh_daily)
# str(kwh_academic)
```

```{r Annual, include = FALSE}
annual_admin <- kwh_annual_df %>%
  filter(type == "Admin")

daily_admin <- kwh_daily %>%
  filter(type == "Admin") %>%
  mutate(date = ymd(date), 
    month = month(date, label = TRUE, abbr = TRUE), 
    day_of_week = wday(date, label = TRUE, abbr = TRUE), 
    kwh_sqft_yr = (kwh * 365) / sqft)

academic_admin <- kwh_academic %>%
  filter(type == "Admin")

ghg_kg <- 0.30082405

cost <- 0.08138507

```

```{r final_admin, include = FALSE}
final_annual_admin <- annual_admin %>%
  mutate(CO2e.MT = round(((kwh_corr*ghg_kg)/1000), digits = 0),
         est.cost = round(kwh_corr*cost),
         kwh_sqft = round(kwh/sqft, digits = 1),
         days_perc = round(days_perc, digits = 0),
         kwh = round(kwh, digits = 0),
         kwh_corr = round(kwh_corr, digits = 0))
# final_annual_admin

final_academic_admin <- academic_admin %>%
   mutate(CO2e.MT = round(((kwh_corr*ghg_kg)/1000), digits = 0),
         est.cost = round(kwh_corr*cost),
         kwh_sqft = round(kwh/sqft, digits = 1),
         days_perc = round(days_perc, digits = 0),
         kwh = round(kwh, digits = 0),
         kwh_corr = round(kwh_corr, digits = 0))
# final_academic_admin1
```

```{r final_admin_2, include = FALSE}
final2_annual_admin <- 
  arrange(final_annual_admin, kwh_corr) %>%
  select(meter, NAME, days_perc, sqft, kwh, kwh_corr, kwh_sqft, est.cost, CO2e.MT)

final2_academic_admin <- 
  arrange(final_academic_admin, kwh_corr) %>%
  select(meter, NAME, days_perc, sqft, kwh, kwh_corr, kwh_sqft, est.cost, CO2e.MT)

# str(final2_academic_admin)
summary(final2_annual_admin)

# str(final2_annual_admin)
summary(final2_annual_admin)

```

## Electricity Use Summary

  Table 1 summarizes electricity use in FY25 for each individually-metered administrative building on the Dickinson College campus. 5 North Orange Street is the largest building with the largest total electricity use and largest electricity use per square foot. 554 West Louther Street is the smallest building with the third smallest total electricity use but the second highest electricity use per square foot. Based on the median values for the category, the ‘typical’ administrative building has the following characteristics:
  
  - **Square footage:** 14,650
  - **kWh/year:** 37,884 kWh/yr
  - **kWh/sqft:** 3.2 kWh/sqft
  - **Cost (*$*):** $3,319
  - **CO2e (*MT*):** 12.5 MT
  
  We see vast variability in electricity use across administrative buildings on campus and the associated costs and greenhouse gas emissions. 5 North Orange Street was by far the largest consumer of total electricity use per fiscal year and per square foot, at 567,745 kWh/year and 8 kWh/sqft respectively, while 556 West Louther Street and Old West were tied for smallest electricity use per square foot, both at about **2 kWh/sqft**. 556 West Louther Street had the smallest total electricity use per fiscal year at 4,985 kWh/sqft. The EIA categorizes administrative or professional office buildings at other Colleges and Universities as using an average of 227,000 kWh per building per year, with a median electricity use per square foot of **9.7 kWh/sqft** (Figures 1 and 2).

```{r Table_1}
datatable(final2_annual_admin,
          rownames = FALSE,
          colnames = c("Meter", "Building", "Days of data (%)", "Square footage", "kWh per year", "kWh (corrected)", "kWh per sqft", "Cost ($)", "CO2e (MT)"),
          filter = "none",
          class = "compact",
          options = list(pageLength = 6, autoWidth = TRUE, dom ='t'), caption =
"Table 1. Total electricity use, estimated financial cost, and estimated greenhouse gas emissions of administrative buildings during fiscal year 2025.")
```

Table 2 illustrates the largest electricity user for the academic year was 5 North Orange Street at **371,381 kWh/year** and **5.0 kWh/sqft** while the smallest electricity user was 556 West Louther Street at **2,791 kWh per year** and **1.4 kWh/sqft**. This is different from the fiscal year where Old West was one of the smallest electricity users per square foot. During the academic year, Old West used the second least electricity per square foot. Dickinson College’s administrative buildings’ electricity use still remained low compared to the administrative buildings on other college campuses compiled by the EIA, both in terms of total electricity use and per square foot.

```{r Table_2}
datatable(final2_academic_admin,
          rownames = FALSE,
          colnames = c("Meter", "Building", "Days of data (%)", "Square footage", "kWh per year", "kWh (corrected)", "kWh per sqft", "Cost ($)", "CO2e (MT)"),
          filter = "none",
          class = "compact",
          options = list(pageLength = 6, autoWidth = TRUE, dom ='t'), caption = "Table 2. Total electricy use, estimated financial cost, and estimated greenhouse gas emissions of administrative buildings during academic year 2024/2025.")
```

Figure 1 shows how much electricity the college’s administrative buildings used each month of the year. Please note the y-axis is the same for each graph, meaning if one building has extremely high electricity use, like 5 North Orange Street does, it will make the other buildings appear as though they use little to no electricity. 5 North Orange Street (as seen from Tables 1 and 2) consistently has much high electricity use than any other building in its category.

```{r Figure_1, warning = FALSE}
ggplot(daily_admin) +
  geom_col(aes(x = month, y = kwh, fill = NAME)) +
  facet_wrap(~ NAME) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), 
    legend.position = "none") +
  labs(title = "Figure 1. Monthly patterns in electricity use for administrative buildings",
       subtitle = "Fiscal Year 2025",
    x = NULL,
    y = "Electricity use (kWh)")
```

Figure 2 shows how the electricity intensity (kWh/sqft/year) for administrative buildings changes throughout the year. The median usage of a building is used to rank it from highest to lowest. The vertical dashed black line shows the national median (*10.1 kWh/sqft/year*), and the shaded gray band shows the interquartile range (*6 to 16 kWh/sqft/year*) from the EIA 2018 Commercial Buildings Energy Consumption Survey.

```{r Figure_2, warning = FALSE}
eia_median <- 10.1
eia_q25 <- 6.1    
eia_q75 <- 15.7   

ggplot(daily_admin, aes(x = reorder(NAME, kwh_sqft_yr, FUN = median, na.rm = TRUE), 
    y = kwh_sqft_yr,
    fill = NAME)) +
  annotate("rect", ymin = eia_q25, ymax = eia_q75, xmin = -Inf, xmax = Inf, 
           alpha = 0.2, fill = "gray50") +
  geom_hline(yintercept = eia_median, linetype = "dashed", color = "black", linewidth = 0.8) +
  geom_boxplot(alpha = 0.9, outlier.alpha = 0.5) +
  coord_flip() +
  facet_wrap(~ period) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(title = "Figure 2. Electricity intensity for administrative buildings",
       subtitle = "Fiscal Year 2025",
    caption = "White dotted line and gray shaded area show median and 25th to 75th percentile of typical values in U.S. office buildings on college campuses (EIA 2022)",
    x = NULL, 
    y = "Electricity Intensity (kWh/sqft/year)")
```

## Recommendations

  Dickinson’s administrative buildings generally operate below the national EIA median for electricity intensity at 10.1 kWh/sqft/year. Despite this overall efficiency, investigating high-use or high-intensity outliers could support targeted energy reductions. For instance, 5 North Orange Street has an electricity usage more than five times higher than the next-largest administrative building. Reducing electricity use in this building would likely result in significant savings.
  
  As mentioned in the Background section, there was only individual data for six of the eleven buildings categorized as administrative on campus. Without the other five buildings, a full picture of administrative buildings on campus is not clear. We recommend implementing submeters on the Weis and Main Meters to better understand our administrative buildings' electricity use through the entirety of campus.  

  To better contextualize these measurements, we recommend conducting targeting comparisons. For example, we suggesting investigating:
  
  - 554 West Louther vs. 556 West Louther: Both are ROTC offices and fairly similar buildings, but 554 W. Louther has an electricity intensity over twice as high. It would be valuable to identify the factors driving this difference. 
  - 5 North Orange Street in the Administrative category vs. the part in the "Other" category: This comparison could point to specific areas of the building that should reduce electricity use, depending on the department's necessary functions. 
  - 50 Mooreland Avenue vs. Old West: This comparison could provide valuable insights into office habits that can be implemented at 50 Mooreland Avenue. While Old West and 50 Mooreland Avenue have similar 9-5 weekday schedules, Old West uses less electricity per square foot than 50 Mooreland Avenue (*1.7 kWh/sqft and 2.5 kWh/sqft, respectively*). Energy saving habits may be implemented in Old West that are not in place at 50 Mooreland Avenue and could be adopted in the future. 
  
  Ultimately, **mapping operating hours, office habits, and localized equipment loads** will provide the necessary context to explain findings that cannot be inferred from meter data alone.

## Sources

Leary, N. (2025). **Dickinson College Greenhouse Gas Inventory**. Dickinson College.

U.S. Energy Information Administration. (2022). **2018 Commercial Buildings Energy Consumption Survey (CBECS)**. U.S. Department of Energy. <https://www.eia.gov/consumption/commercial/>.

## Statement of AI Use

Amiya used AI to create Figures 1 and 2. Eva did not use AI to create Tables 1 and 2. No writing was created using AI.

## Partner Contributions

Amiya Marbles drafted the Background section, Electricity Use Summary section, and Figures 1 and 2 (including data preparation). Eva Hanlin created Tables 1 and 2 (including data preparation) and wrote the Recommendations section. Both partners contributed to revisions.
